---
title: "PCA-demo"
author: "Ethan Weed"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
library(tidyverse)

data(iris)
df <- iris
```
## Calculate PCA

Here are three different functions for calculating PCA. ``princomp()`` uses the Spectral decomposition method, which looks at correlations/covariance between variables. ``factoextra::PCA`` and ``prcomp`` use the # factoextra PCA Singular value decomposition method: looks at correlations/covariance between individuals.

### Method 1: princomp
```{r}
m1 <- princomp(df[,1:4])
```

### Method 2: FactoMineR
```{r}
library(FactoMineR)
m2 <- PCA(df[,1:4], scale.unit = TRUE)
```

### Method 3: prcomp
```{r}
m3 <- prcomp(df[,1:4], scale = TRUE)
```

## Extract PCA coordinates

### using FactoMineR
```{r}
library(FactoMineR)
# For variable coordinates
#VarM1 <- get_pca_var(m1)
#VarM2 <- get_pca_var(m2)
#VarM3 <- get_pca_var(m3)

# For observation coordinates
VarM1 <- get_pca_ind(m1)
VarM2 <- get_pca_ind(m2)
VarM3 <- get_pca_ind(m3)

VarM1 <- VarM1$coord
VarM2 <- VarM2$coord
VarM3 <- VarM3$coord

colnames(VarM1) <- c("PC1", "PC2", "PC3", "PC4")
colnames(VarM2) <- c("PC1", "PC2", "PC3", "PC4")
colnames(VarM3) <- c("PC1", "PC2", "PC3", "PC4")
```

### extracted directly from pca objects
```{r}
# Method 1: princomp
m1_coordinates <- data.frame(m1$scores)

# Method 2: factoextra
m2_coordinates <- data.frame(m2$ind$coord)

# Method 3: prcomp
m3_coordinates <- data.frame(m3$x)
```


# Compare the first PC as calculated by methods 1,2 and 3: do different algorithms give different results?
```{r}
PC1s <-data.frame(m1_coordinates$Comp.1, 
                  m2_coordinates$Dim.1, 
                  m3_coordinates$PC1)

colnames(PC1s) <- c("princomp", "factoextra", "prcomp")
```


# Nice visualizations you can do with FactoMineR

## Scree plot: % variance explained by PC's
```{r}
fviz_eig(m3)
```


## Plot individuals on PC's 1 and 2 and color by category
```{r}
# Add the categorical ("qualitative") data back in from column 5 (Species)
pc <- PCA(df, quali.sup = 5)

# plot individuals on PC's 1 and 2 and color by category
fviz_pca_ind(pc,
             col.ind = iris$Species)
```

## Define overlapping or non-overlapping zones for categories
```{r}
fviz_pca_ind(pc,
             col.ind = iris$Species,
             addEllipses = TRUE)
```


## Show the contribution of the predictor variables to PC's 1 and 2
```{r}
fviz_pca_var(pc,
             col.var = "contrib")
```

