---
title: "Partial correlation"
author: "Ethan Weed"
date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, ppcor, pander)
```

According to <a href="https://en.wikipedia.org/wiki/Partial_correlation" target="_blank">Wikipedia</a>, partial correlation "measures the degree of association between two random variables, with the effect of a set of controlling random variables removed." To make things simple, we take three variables from the mtcars dataset.

```{r}
d <- mtcars

df <- d[,c("mpg", "drat", "wt")]

```

To compute the partial correlation, we 

1. model two linear regressions with the same predictor variable and different outcome variables
2. correlate the resisduals from the two models

```{r}
mod1 <-  lm(drat ~ wt, data = d)
m1s <- summary(mod1)
m1res <- m1s$residuals

mod2 <-  lm(mpg ~ wt, data = d)
m2s <- summary(mod2)
m2res <- m2s$residuals

paste("Partial correlation of drat and mpg is ", cor(m1res,m2res))
```

To check, we can compare with the result using the ppcor package:

```{r}
pander(data.frame(pcor(df))[,1:3])
```

As expected, the partial correlation between drat and mpg is 0.1806, just as we found doing it the long way. 

The same process works for more than three variables. Here, we model drat predicted by wt + qsec, and then mpg predicted by wt + qsec. We find the residuals of both regression models, and correlate them. The result is the same as that given by ppcor.

```{r}

df <- d[,c("mpg", "drat", "wt", "qsec")]

mod1 <-  lm(drat ~ wt + qsec, data = d)
m1s <- summary(mod1)
m1res <- m1s$residuals

mod2 <-  lm(mpg ~ wt + qsec, data = d)
m2s <- summary(mod2)
m2res <- m2s$residuals

paste("Partial correlation of drat and mpg is ", cor(m1res,m2res))

pander(data.frame(pcor(df))[,1:4])
```



Thus, a partial correlation matrix is a matrix of the residual correlations between variables, when the effects of the remaining variables have been removed.
